The Etiology of Cybersecurity

### 2020-09-03 reports/paper_0/main.tex

If we consider a functional block that generates nonces (i.e. number used
only once), the semantics of this block must define a procedure such that 
a freshness requirement holds. By the RIDI-Hypothesis we have that
the following weaknesses may occur:
\begin{itemize}
	\item Replace: the output-nonces of the block are replaced with
		non-fresh numbers
	\item Insert: the output-nonces are concatenated with other numbers,
		resulting in non-fresh nonces. For example, the nonces 110 and
		111 are modified by inserting a 1 and a 0 at the beginning and
		end respectively, resulting in the two identical (and the
		non-fresh) nonces 1110 and 1110
	\item Delete: part of output-nonce is dropped. For example, the last bit of 1101 and 1100 is dropped, resulting in the two equal and non-fresh nonces: 110 and 110.
	\item Inject: substitution at bit level results in non-fresh nonces. For example, 110 becomes 111 becomes equal by flipping either the last 0 of the former nonce or the last 1 of the latter.
\end{itemize}

@@

\subsection{Security and Insecurity of a System}
\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
	A system security design is given by a precise system 
	specification over the physical and functional architectures, that
	uniquely, i.e.  tested against the range of design possible in the
	$\abftheory$-framework, defines the design built on top of those
	requirements.
\end{hypothesis}

@@

The tool generates a graph-like internal structure which represents
the specification (called $\abftheory$-graph). The $\abftheory$-graph defines the system as
a number of regions of assertions, beliefs, and facts. Those regions
are connected by a generic relation which is evaluated as follows.
The graph is translated into 
a logical formula that represents the specification in the $\abftheory$-framework and,
along with the axiomatization of the RCC5 calculus, is given as input to the Z3 SMT solver.
The solver identifies all possible configurations of the system and, in turn,
identifies all potential weaknesses. 
The internal structures of the tool
can be viewed as PDF and the results are reported into an Excel file.
The Excel file also reports the total number of configurations as indicating
the cybersecurity risk associated to the specification. A user can change the status
of each weakness in the Excel file, from the initial (default) status (open)
to ``mitigated''. As a consequence, the risk is re-calculated on-the-fly, i.e.
without the need of running the tool again, based on annotations and formulas 
in the Excel file.
In our approach, cybersecurity requirements are not imposed by the
specification but are automatically extracted by our tool as mitigations to potential weaknesses.
Those weaknesses are related to the different insecure configurations 
of the specified system.

\paragraph{Case Studies}
We report the results of the evaluation of a water level reader (sensor ad-hoc example). 
\begin{figure}
	\centering
	\includegraphics[width=.5\columnwidth]{eng_cs1.png}
	\caption{Water level reader deployment diagram}
	\label{fig:eng_cs1}
\end{figure}
As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
which represent the physical reader that needs to be placed in a tank, and 
the board that interprets the readings and outputs them as digital signals.
The two components are connected by a wire.
\begin{figure}
	\centering
	\includegraphics[width=.9\columnwidth]{internal_cs1.png}
	\caption{Sensor Board Object Diagram}
	\label{fig:int_cs1}
\end{figure}
In Figure~\ref{fig:int_cs1}, we report the functional architecture that
receives the incoming communications from the sensor in the tank, and 
communicates them encrypted.
\begin{figure*}
	\centering
	\includegraphics[width=.7\textwidth]{results_excel.pdf}
	\caption{Partial View of the results in the Excel file}
	\label{fig:results}
\end{figure*}
The results, summarized in Figure~\ref{fig:results}, reports 16777216
possible scenarios in which at least one component diverge from the specification.

@@

\section{Conclusion and Future Works}
%\fixnote{mv}{Se ho capito bene, detto a parole mie, quello che il vostro
%approccio permette di fare, dati requisiti e architettura del sistema, è di
%elencare in maniera sistematica tutti i "punti" in cui potrebbero sorgere dei
%problemi. Spiegherei un po' meglio, qui nelle conclusioni o nella sezione
%precedente, come questo approccio dovrebbe essere usato nella pratica... Voglio
%dire che non mi sembra necessariamente in opposizione ad altri approcci ma
%magari può integrarsi a questi.  Inoltre quando parlate dei passi successivi,
%ad esempio il considerare la generazione di test cases, se avete già qualche
%idea su come procedere, potreste discuterne brevemente.}
We proposed an hypothesis for a foundational theory on security, arguing that
cybersecurity-related issues are not related to the maliciousness of an agent but to
the vagueness of the cybersecurity controls on the engineering processes.
We also provided a prototype tool for the quantitative estimation of the cybersecurity
risk, based on a UML model of the system under assessment.
%The current state-of-the-art processes allow, given a specification,
%an engineer to design the system in such a way that cybersecurity issues arise due to the lack
%of proper cybersecurity risk assessment processes.
%Those designs, again, lack of cybersecurity verification processes based on a solid 
%cybersecurity foundational theory and then permit the generation of insecure implementation.
The cybersecurity verification and test-case generation will be the focus of our next steps.
%We conclude that the problem of security is a problem related to the 
%many possible design and, in turn, implementation given a specification.
%The problem is related to the epistemological search for truth,
%where the challenge is to relate information and belief to human knowledge.
%Scientifically, generalizing the human to an agent, the problem is
%to relate assertions and beliefs, to facts. From an engineering standpoint,
%by generalizing the concept of agents as architectural subsystems, the
%problem is to link channels (and ports) and functional architectures to 
%requirements.

>>>>

If we consider a functional block that generates nonces (i.e. number used
only once), the semantics of this block must define a procedure such that 
a freshness requirement holds. By the RIDI-Hypothesis we have that
the following weaknesses may occur:
\begin{itemize}
	\item Replace: the output-nonces of the block are replaced with
		non-fresh numbers
	\item Insert: the output-nonces are concatenated with other numbers,
		resulting in non-fresh nonces. For example, the nonces 110 and
		111 are modified by inserting a 1 and a 0 at the beginning and
		end respectively, resulting in the two identical (and the
		non-fresh) nonces 1110 and 1110
	\item Delete: part of output-nonce is dropped. For example, the last bit of 1101 and 1100 is dropped, resulting in the two equal and non-fresh nonces: 110 and 110.
	\item Inject: substitution at bit level results in non-fresh nonces. For example, 110 and 111 becomes equal by flipping either the last 0 of the former or the last 1 of the latter.
\end{itemize}

@@

\subsection{Security and Insecurity of a System}\label{sec:formula}
\begin{hypothesis}{\bf System Security Design --}\label{hyp:security}
	A system security design is given by a precise system 
	specification over the physical and functional architectures, that
	uniquely, i.e.  tested against the range of design possible in the
	$\abftheory$-framework, defines the design built on top of those
	requirements.
\end{hypothesis}

@@

The tool generates a graph-like internal structure which represents
the specification (called $\abftheory$-graph). The $\abftheory$-graph defines the system as
a number of regions of assertions, beliefs, and facts. Those regions
are connected by a generic relation which is evaluated as follows (according to
the formula in Section~\ref{sec:formula}).
The graph is translated into 
a logical formula that represents the specification in the $\abftheory$-framework and,
along with the axiomatization of the RCC5 calculus, is given as input to the Z3 SMT solver.
The solver identifies all possible configurations of the system and, in turn,
identifies all potential weaknesses. 
The internal structures of the tool
can be viewed as PDF and the results are reported into an Excel file.
The Excel file also reports the total number of configurations as indicating
the cybersecurity risk associated to the specification. A user can change the status
of each weakness in the Excel file, from the initial (default) status (open)
to ``mitigated''. As a consequence, the risk is re-calculated on-the-fly, i.e.
without the need of running the tool again, based on annotations and formulas 
in the Excel file.
In our approach, cybersecurity requirements are not imposed by the
specification but are automatically extracted by our tool as mitigations to potential weaknesses.
Those weaknesses are related to the different insecure configurations 
of the specified system.

\paragraph{Case Studies}
We report the results of the evaluation of a water level reader (sensor ad-hoc example). 
\begin{figure}
	\centering
	\includegraphics[width=.5\columnwidth]{eng_cs1.png}
	\caption{Water level reader deployment diagram}
	\label{fig:eng_cs1}
\end{figure}
As in Figure~\ref{fig:eng_cs1}, we defined 2 agents: sensorInTank and sensorBoard,
which represent the physical reader that needs to be placed in a tank, and 
the board that interprets the readings and outputs them as signals.
The two components are connected by a wire.
\begin{figure}
	\centering
	\includegraphics[width=.9\columnwidth]{internal_cs1.png}
	\caption{Sensor Board Object Diagram}
	\label{fig:int_cs1}
\end{figure}
In Figure~\ref{fig:int_cs1}, we report the functional architecture that
receives the incoming communications from the sensor in the tank, and 
communicates them encrypted.
\begin{figure*}
	\centering
	\includegraphics[width=.7\textwidth]{results_excel.pdf}
	\caption{Partial View of the results in the Excel file}
	\label{fig:results}
\end{figure*}
The results, summarized in Figure~\ref{fig:results}, reports 16777216
possible scenarios in which at least one component diverge from the specification.

@@

\section{Conclusion and Future Works}
%\fixnote{mv}{Se ho capito bene, detto a parole mie, quello che il vostro
%approccio permette di fare, dati requisiti e architettura del sistema, è di
%elencare in maniera sistematica tutti i "punti" in cui potrebbero sorgere dei
%problemi. Spiegherei un po' meglio, qui nelle conclusioni o nella sezione
%precedente, come questo approccio dovrebbe essere usato nella pratica... Voglio
%dire che non mi sembra necessariamente in opposizione ad altri approcci ma
%magari può integrarsi a questi.  Inoltre quando parlate dei passi successivi,
%ad esempio il considerare la generazione di test cases, se avete già qualche
%idea su come procedere, potreste discuterne brevemente.}
We proposed an hypothesis for a foundational theory on security, arguing that
cybersecurity-related issues are not linked to the maliciousness of an agent but to
the vagueness of the cybersecurity controls on the engineering processes.
We also provided a prototype tool for the quantitative estimation of the cybersecurity
risk, based on a UML model of the system under assessment.
%The current state-of-the-art processes allow, given a specification,
%an engineer to design the system in such a way that cybersecurity issues arise due to the lack
%of proper cybersecurity risk assessment processes.
%Those designs, again, lack of cybersecurity verification processes based on a solid 
%cybersecurity foundational theory and then permit the generation of insecure implementation.
The cybersecurity verification and test-case generation will be the focus of our next steps.
%We conclude that the problem of security is a problem related to the 
%many possible design and, in turn, implementation given a specification.
%The problem is related to the epistemological search for truth,
%where the challenge is to relate information and belief to human knowledge.
%Scientifically, generalizing the human to an agent, the problem is
%to relate assertions and beliefs, to facts. From an engineering standpoint,
%by generalizing the concept of agents as architectural subsystems, the
%problem is to link channels (and ports) and functional architectures to 
%requirements.
